\chapter{Colophon}

{\scshape This is my first} attempt to try a real big document in
\LaTeX. Failed to digest \textit{Knuth}, I had to swallow
\textit{Lamport}. And I loved it. \LaTeX{} made typing this thesis
a pleasure and in the process I started gnawing a bit on \TeX{}
and hope someday I'll write in it.

The citations are in \textsf{\small apalike} style. I hacked the
original \textsf{\small bst} file to create some modifications
that I wanted. The text font is Adobe Minion and the title font
Adobe Myriad. These type1 fonts were imported into \LaTeX{} by
\textsf{\small fontinst} package. I generated the Bib\TeX{}
\textsf{\small bib} files directly from Pubmed citations by a perl
script that I wrote. The figures are made in Macromedia Freehand
10 and exported as PDF before importing into PDF\TeX. Plasmid maps
are drawn by an web-based application Savvy
({\small\url{http://www.bioinformatics.org/savvy}}) that I wrote
as part of my \textsf{\small BioSVG} project. Maps drawn by Savvy
are in Scalable Vector Graphics (SVG) format. The chromatogram
files are also drawn by my own \textsf{\small ABITrace} perl
module in SVG format. I converted the SVG file to PDF by Apache
project's FOP into PDF before importing them into PDF\TeX{}. All
the plots in this thesis were drawn with GNUPlot, another
marvellous gift from free software community.

The pictures on the chapter separators are illustrations by John
Tenniel in Modern Library's \e{The Complete Works of Lewis
Carroll}.

Lastly, I proudly declare this thesis M\$ Office free. None of
Micro\$oft's bloated and buggy code was used to make any part of
this thesis. At last I did something \emph{unofficial}. \ding{45}
